---
title: "Avian  Database Manual v4"
author: "Someone, someone"
date: ""
fontsize: 12pt
header-includes:
   - \usepackage{booktabs}
   - \usepackage{dcolumn}
   - \usepackage{wrapfig}
   - \usepackage{subcaption}
   - \usepackage{caption}
   - \usepackage[font=small,labelfont=bf]{caption}
   - \hypersetup{colorlinks=false}
   - \usepackage{longtable}
   - \usepackage{leading}
   - \leading{13pt}
   - \usepackage{float} #use the 'float' package
   - \floatplacement{figure}{H} #make every figure with caption = h
documentclass: "article"
output:
  pdf_document:
    fig_caption: yes
    #keep_tex: yes
    fig_height: 6
    fig_width: 7
    latex_engine: pdflatex
    number_sections: true
    toc: yes
    toc_depth: 4
    citation_package: "natbib"
linestretch: 2
mainfont: "Linux Libertine O"
bibliography: ThesisLibrary.bib
csl: american-political-science-association.csl
biblio-style: apalike
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos="h")

```

# INTRODUCTION

The boreal region of Canada hosts one of the most diverse bird communities in North America, comprised
of more than 300 species. During the summer, more than 300 bird species and up to three billion birds breed
in the region, leading to its characterization as a North American bird nursery. Historically, these forests have
been subject to little widespread development pressure, and access has been extremely limited, resulting in a
corresponding gap in research and monitoring. However, pressures are mounting, and with the rapid development
of the boreal forest that is presently occurring, there is an urgent need to understand the impact of changing
habitats on boreal bird populations in order to inform management actions. There has been a recent surge
of boreal bird studies across Canada, thus it is timely to undertake a synthesis of existing research, and as a first step towards scenario analysis and decision-support, formulate habitat-based predictive models of species
abundance and distribution for Canadian boreal forests. We envision a series of spatially-explicit, bird-habitat
models, broadly accessible to all organizations interested in boreal conservation planning, and updated regularly
with monitoring data and new research results. The objective of the BAM Avian DB is to assemble and harmonize existing data sets on all boreal forest birds and their habitats through cooperative efforts with boreal bird researchers and associated agencies.

# BAM AVIAN DATABASE
The Boreal Avian Modelling (BAM) Project has been assembling a comprehensive database of systematic, spatially referenced, observational data on boreal birds that have been collected in the boreal and hemiboreal regions of Canada and the United States. Our primary focus is point-count data. This standardized survey technique, used widely in avian research, is well suited for evaluating habitat requirements of forest songbirds.

To date, we have compiled much of the point-count data collected in the Canadian boreal forest since 1993. The avian data set contains information from more than 105 individual research projects comprising more than 100 000 off-road point-count locations, and over 1 million bird records. Approximately 100 bird species are represented. Recent data compilation efforts have resulted in the acquisition of the majority of geo-referenced point-count survey data from across boreal and Arctic Alaska. Efforts continue to acquire data from the hemiboreal region across North America including the Upper Midwest and New England areas of the US.

## Database Structure

The Avian Database contains the observational survey data contributed from all projects, standardized into a common format and ready to use in analyses. The observational data and covariates are maintained in a relational database. It consists of 4 main tables: 

1. Project Summary Table – Preserves essential metadata for each project, including sampling protocol employed, and any restrictions on data use;
2. XY Table – value that pertain geographic coordinates for each survey location;
3. PKEY Table – values that pertain to data and time of visit to a particular survey location, as well as observer identity if available and included; 
4. Point Count Table – values that pertain the actual survey information, such as the number of individuals of each species detected, any information about distance to the detected species, the time of detection relative to survey initiation, and available information on the bird’s behavior.

The main body of this report defines each of the four tables and tabulates the attributes and their characteristics. A summary of the data structure and data dictionary is also presented for each table. Maintaining this information within the database ensures consistent classifications across users, and prevents repetition of background searches on life history information each time a new analysis is undertaken.

## Project Summary Table

The BAM database contains contributed data from 135 distinct projects. Projects are sets of data contributed by a single owner or entity. In principle, all data within a single contributed project were collected under a common survey protocol. The two most important elemenents of a survey protocol are the sampling duration (DURMET) (e.g. 5, 10) and the sampling distance or radius (DISTMET) (e.g. 50m, 100m, unlimited distance) (Tanle 1). Other components of a sampling protocol include the number of years over which the study was conducted,and the number of visits per year. These characteristics are not always uniform for all surveyed locations within a project. The field Methods allow to distinguish within a project difference in sampling duration and distance.   

```{r, echo = FALSE}
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)

wfolder <-getwd()
csv_ps <- file.path(wfolder, "/data/tbl_ProjectSummary2.csv")
tbl_ps <- read.csv(file = csv_ps, sep = ",", check.names=FALSE)

tblps_title <- "Project Summary Table Structure"
kable(tbl_ps, "latex", booktabs = T, caption = tblps_title, linesep = "", row.names = FALSE) %>%
  kable_styling(font_size = 9, latex_options= "hold_position")  %>%
  column_spec(2, width = "20em")
```

### PCODE
Each Project is assigned a unique code (PCODE) which links to data sharing agreements, and also serves as a key to relate project information to stations (Table 2). A single project can however use different methodology. To link project information to observations found in PointCount Table, use Method field. 

```{r, echo = FALSE}
cat("\n")

wfolder <-getwd()
csv_01 <- file.path(wfolder, "/data/table01.csv")
tbl_01 <- read.csv(file = csv_01, sep = ",", check.names=FALSE)

tbl01_title <- "List of PCODE used in BAM Avian DB v4"
kable(tbl_01, "latex", longtable = T, booktabs = T, caption = tbl01_title, linesep = "", row.names = FALSE) %>%
  kable_styling(font_size = 9, latex_options = c("repeat_header"))

```


### Method
This field is the unique project identification code. It is the concatenation of PCODE with a unique identifier. It allows to link Project metadata to Point Count data according to the methodology used during the survey.

### Durmeth
Durmeth is the duration methodology used during the survey. Within these durations, observations were sometimes recorded by interval class, as between 0-3 and 3-5 minutes within a 5 minute point count survey. The unit is in minutes (Table 3).


```{r, echo = FALSE}
wfolder <-getwd()
csv_durMeth <- file.path(wfolder, "/data/tbl_DurationMethodology.csv")
tbl_durMeth <- read.csv(file = csv_durMeth, sep = ",", check.names=FALSE)

tbl_durMeth_title <- "List of Duration Methodology Code"
kable(tbl_durMeth, "latex", booktabs = T, caption = tbl_durMeth_title, linesep = "", row.names = FALSE, align = c('c','l')) %>%
  kable_styling(font_size = 9, latex_options= "hold_position")
```


### Distmeth
Dismeth is the distance methodology over which data were recorded during the survey. It is also called detection radius. Some protocols stratified observations by interval class, such as less than 50m, 50-100m and greater than 100m. Distance unit is in meters (Table 4).

```{r, echo = FALSE}
wfolder <-getwd()
csv_distMeth <- file.path(wfolder, "/data/tbl_DistanceMethodology.csv")
tbl_distMeth <- read.csv(file = csv_distMeth, sep = ",", check.names=FALSE)

tbl_distMeth_title <- "List of Distance Methodology Code"
kable(tbl_distMeth, "latex", longtable = T, booktabs = T, caption = tbl_distMeth_title, linesep = "", row.names = FALSE, align = c('c','l')) %>%
  kable_styling(font_size = 9, latex_options= c("repeat_header"))
```

### Changemethod
Changemethod indicates if the methodology change at any point of the project (over years or within year). Most of the time, changes will be reflected by a change in Method field.

```{r, echo = FALSE}
wfolder <-getwd()
csv_changeMeth <- file.path(wfolder, "/data/tbl_yesno.csv")
tbl_changeMeth <- read.csv(file = csv_changeMeth, sep = ",", check.names=FALSE)

tbl_changeMeth_title <- "List of Change Method Code"
kable(tbl_changeMeth, "latex", booktabs = T, caption = tbl_changeMeth_title, linesep = "", row.names = FALSE, align = c('c','l')) %>%
  kable_styling(font_size = 9, latex_options= "hold_position")
```

### Project
Project identifies the project name from which the point count data came.

### Partner
Partner identifies the name of the contributor.

### Site_fix
Site-fix was specifically created for Diana's analysis. Allow to reassign sites for project where they were too dispersed. 

```{r, echo = FALSE}
wfolder <-getwd()
csv_changeMeth <- file.path(wfolder, "/data/tbl_yesno.csv")
tbl_changeMeth <- read.csv(file = csv_changeMeth, sep = ",", check.names=FALSE)

tbl_changeMeth_title <- "List of Change Method Code"
kable(tbl_changeMeth, "latex", booktabs = T, caption = tbl_changeMeth_title, linesep = "", row.names = FALSE, align = c('c','l')) %>%
  kable_styling(font_size = 9, latex_options= "hold_position")
```

### Years
Years indicates number of year sampled. It does not imply that each station was visited over repeated years.

### Rounds
Rounds indicates number of rounds sampled in a year. 

### Maxduration
Maxduration represents the maximum time of individual count. It is directly derived from dur.Meth

### Intervals
Intervals is the number of time intervals used in count. For example, if the duration sampled is 0-3min, 3-5min, number of intervals  = 2. 

### Maxdist
Maxdist represents the maximum distance sampled of individual count in the project methodology. It is directly derived from distMeth. 

### Bands
Bands represents the number of distance bands in sampling method. For example, if bird count were recorded within a radius classes 0-25m,  25-50m, 5-100m, bands = 3. 

### Use_restricted
Use_restricted identifies whether the use of the data is unrestricted, restricted or limited.

### BBS
BBS identifies projects where protocols are BBS specific but not bbs data.

### Origin
The project summary table was built after the main database - Origin was to track where the bulk of a dataset was being pulled from - also the Atlas data is a bit separate, the ALMS data is a huge amount of data from Alaska

### Protocol
Protocol identifies if projects protocols was derived from another project (i.e.; ATLAS).



## XY Table 
Stations are the geographically referenced locations where individual point counts were conducted. Each station
is linked to a Project via (PCODE) ,and the location is recorded in geographic coordinates. A unique key
(SS)is assigned to each station. In constructing the key, we used as far as possible the project-specific
naming conventions; this allows us preserve some features of the sampling designs for individual projects,
and to identify groups of stations that should potentially be grouped in random effects models. We allow
one level of grouping within projects, and stations are numbered within groups. There is also a unique
numeric identifier (ID_link) assigned to each station, independent of the constructed key. This table may
also include one or more records spatially locating the station within ecoregions, BCRs or other geographic
zonations that can be coded as factors. However, such attributes are not guaranteed to be maintained
across database versions, and should properly be located in a separate table linked by (SS).


```{r, echo = FALSE}
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)

wfolder <-getwd()
csv_ps <- file.path(wfolder, "/data/tbl_XY.csv")
tbl_ps <- read.csv(file = csv_ps, sep = ",", check.names=FALSE)

tblps_title <- "XY Table Structure"
kable(tbl_ps, "latex", booktabs = T, caption = tblps_title, linesep = "", row.names = FALSE) %>%
  kable_styling(font_size = 9,  latex_options= "hold_position")
```

###PCODE 
###SS 
###X  
###Y  
###Timezone  
###NRCANlink 
###TYPE 
###OnRoad 
###SS_SITE2  
###BOR_LOC 
###Resample 

  
## PKEY Table 

Rounds record the year, calendar date and clock time when data were collected at a station. Clock time is
time-zone corrected given the location and date of sampling. For stations with multiple visits within years,
a Round number is also recorded. Dates are also recorded as Julian dates and relative to the start of the
local growing season, as determined by custom interpolated climate data. Times are also recorded relative
to local nautical(?) sunrise at the geographic coordinates and date of sampling. Each Visit is assigned a
unique PKEY composed of SS, a two-character code for sampling year (e.g. 99 = 1999, 07 = 2007) and the
Round.

## Point Count Table  

Point counts record the actual observations. An observation is defined by a species code and an abundance. Standard
OAU codes are used for species. In addition, each observation is assigned a distance class, a interval class,
and a behavioural code. The distance classes correspond to the distance classes recognised in the Project
protocol. Similarly, the interval class records the time interval during the point count during which the
observation was made. The behavioural code is the recorded bahaviour by which the individual bird or
birds were detected (e.g. countersinging, 
ying overhead). These codes and classes are described in detail
below.

## 1.2 Project level summaries

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
