---
title: "Batch Formatting Geonetwork"
author: "M. Houle"
date: "January 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This script intend to create metadata using Geonetwork standard when dataset have the same format, such as species distribution for individual birds species. 

First, metadata needs to be created inside Geonetwork using string to replace where species name and abbreviation are used. Then, export metadata as a zip file in a work folder. Metadata xml is then process by the script to create new metadata that include species name,  abbrevation, ... and fix link to download to access respective files.  

```{r pressure, echo=FALSE} 

```

```{r Batch upload Species Distribution dataset }  
library(XML)
library(utils)

#List birds table 
#birdtbl <- ("../data/birdAcronyme.csv")
parentDir <-dirname(getwd())
birdtbl <-file.path(parentDir,"data/birdAcronyme.csv")
bird <- read.csv(birdtbl, sep=",", header=TRUE,stringsAsFactors=FALSE)

gnDir <- file.path(parentDir,"GNmetadata/templateSpeciesDistribution/data/metadata")
con <- file(paste0(gnDir,"/metadata.xml"))
gnVersion <-readLines(con, warn=FALSE)

xml_update <- apply(bird,1,function(x) {
                    dirTo <- "/home/ubuntu/geonetwork/toGN/metadata/SpeciesDistribution"
                    species_updated <-gsub("XXXX",x[2],gnVersion)
                    xsname_updated <-gsub("SSSS",x[3],species_updated)
                    abbrev_updated <-gsub("AAAA",x[1],xsname_updated)
                    birdfl<-file(paste0(dirTo,"/",x[1],".xml"))
                    writeLines(abbrev_updated, birdfl)
                    close(birdfl)
})
close(con)

```
